---
title: "How To: PCA Loadings Plot"
output: html_notebook
---
  
```{r}
library(tidyverse)
library(MASS) # for Boston data set
```

Next we seek to reduce the dimension of our predictor space by applying Principal Component Analysis (PCA).  PCA uses linear combinations of possibly correlated input variables to form a new set of variables (principal components) that are uncorrelated with one another.  Additionally, the variables are created sequentially to explain the most variance possible in the dataset.  A variance threshold can be used for dimensionality reduction (e.g. keep only those components that explain more than 5\% of the variance). Note: we center and scale the data before applying PCA since the PCA algorithm uses a distance metric.

Next we consider a loading plot which shows how strongly each input variable influences a primary component.  The length of the vector along the PCX axis (i.e. the length of the vector projected to the PCX axis) indicates how much weight that variable has on PCX.  The angles between vectors tell us how the variables are correlated with one another.  If two vectors are close, the variables they represent are positively correlated.  If the angle is closer to 90 degrees, they are not likely to be correlated.  If the angle is close to 180 degress they are negatively correlated.

```{r}
get_pca_loadings_plot <- function(df, pc_a, pc_b, center = TRUE, scale = TRUE) {
  
  pca_result <- prcomp(~ ., data = df, center = center, scale = scale)
  
  loadings_df <- data.frame(variable = rownames(pca_results$rotation), pca_results$rotation)
  
  wrapr::let(
    alias = list(PCOMP_A = pc_a, PCOMP_B = pc_b),
    expr = {
      ggplot(loadings_df) +
        geom_segment(
          aes(x = 0, y = 0, xend = PCOMP_A, yend = PCOMP_B), 
          arrow = arrow(length = unit(1/2, "picas")),
          color = "black"
        ) +
        annotate(
          "text", 
          x = loadings_df$PCOMP_A*1.2, y = loadings_df$PCOMP_B*1.2,
          label = loadings_df$variable
        ) +
        labs(x = pc_a, y = pc_b) +
        theme_bw()
    }
  )
}

get_pca_loadings_plot(Boston, "PC1", "PC2")
```





