---
title: "How To: PCA Proportion of Explained Variance"
output: html_notebook
---

http://strata.uga.edu/8370/lecturenotes/principalComponents.html

```{r}
library(tidyverse)
library(patchwork) # for wrap_plots()
library(MASS) # for Boston data set
```

Next we seek to reduce the dimension of our predictor space by applying Principal Component Analysis (PCA).  PCA uses linear combinations of possibly correlated input variables to form a new set of variables (principal components) that are uncorrelated with one another.  Additionally, the variables are created sequentially to explain the most variance possible in the dataset.  A variance threshold can be used for dimensionality reduction (e.g. keep only those components that explain more than 5\% of the variance). Note: we center and scale (standardize) the data before applying PCA since variables with larger mean or standard deviation will be prioritized to explain the variation of the data.

In the scree plot below, the magnitude of the eigenvalue indicates the amount of variation that each principal component captures.  The proportion of variance for a given component is the component's eigenvalue divided by the sum of all eigenvalues.  We see that the first component  and second components explain ~35\%  and ~13\% of the variance respectively. Later components are similar in the amount of variance they explain.  The bottom plot shows the cumulative variance explained by the first N components.  One rule of thumb is to keep enough components so that this cumulative variance exceeds 80\%.  In this case, 7 variables appears to be sufficient.

```{r}
get_pca_variance_plots <- function(df, center = TRUE, scale = TRUE, cum_hline = NULL) {
  
  var_proportions <- prcomp(~ ., data = df, center = center, scale = scale) %>%
    summary() %>%
    `[[`("importance") %>%
    t() %>%
    as.data.frame() %>%
    mutate(component_number = str_replace(rownames(.), "PC", "")) %>%
    mutate(component_number = as.numeric(component_number)) %>%
    rename_all(~tolower(str_replace_all(., "\\s+", "_"))) 
  
  scree_plot <- ggplot(var_proportion, aes(x = component_number, y = proportion_of_variance)) +
    geom_col() +
    geom_line() +
    geom_point()
  
  cum_prop_plot <- ggplot(var_proportion, aes(x = component_number, y = cumulative_proportion)) +
    geom_line() +
    geom_point() + 
    geom_hline(yintercept = cum_hline)
  
  list(scree_plot, cum_prop_plot)

}

get_pca_variance_plots(Boston, cum_hline = 0.9) %>%
  wrap_plots()

```


Apply PCA:
```{r}
pca_results <- prcomp(~ ., data = Boston, center = TRUE, scale = TRUE)
```

Get importance measures:
```{r}
importance <- pca_results %>%
  summary() %>%
  `[[`("importance") %>%
  t() %>%
  as.data.frame()
```

Make component numbers numeric instead of "PCX"; make column names in_this_form:
```{r}
importance_clean <- importance %>% 
  mutate(component_number = str_replace(rownames(.), "PC", "")) %>%
  mutate(component_number = as.numeric(component_number)) %>%
  rename_all(~tolower(str_replace_all(., "\\s+", "_"))) 
```

A "scree" plot is a bar chart of components and the variance they explain:
```{r}
importance_clean %>% 
  ggplot(aes(x = component_number, y = proportion_of_variance)) +
  geom_col() +
  geom_line() + 
  geom_point()
```

A plot of cumulative proportion of variance will allow the user to see how many components to keep to explain X amount of variance:
```{r}
importance_clean %>% 
  ggplot(aes(x = component_number, y = cumulative_proportion)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0.9)
```


