---
title: "exploratory_analysis"
author: "ZackBarry"
date: "1/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DataExplorer)

dataset <- read_csv("Data/AB_NYC_2019.csv")
```

## Abstract



## Data source

This data set has nearly 50,000 observations of 16 different variables, 6 of 
which are discrete and 10 of which are continuous.

## Acquiring/Loading Data

```{r}
dataset <- read_csv("Data/AB_NYC_2019.csv")
head(datas)
```

```{r}
# print size of data set
dim(dataset)
```

```{r}
# high-level overview of data types and missing data
plot_intro(dataset)
```

A high-level overview of the types of variables / amount of missing data shows 
that we have 10 continuous and 6 discrete variables.  Most rows have no missing
values, and the total number of missing observations is quite low.  Since
there are 50,000 observations to draw from, we should be able to generate some
great insights!

## Data Wrangling



```{r}
# check which data is missing
plot_missing(dataset)
```

```{r}
# further investigate missing data
dataset %>%
  filter(is.na(reviews_per_month) | is.na(last_review)) %>%
  count(number_of_reviews)
```

Missing observations are mostly limited to `reviews_per_month` and `last_review`.
Upon further investigation, these missing observations occur when `number_of_reviews`
is equal to 0.  This makes sense because if a property has never received any
reviews there is no concept of "most recent review" nor can a rate of reviews be
calculated.  However, it makes sense to set `reviews_per_month` to 0 when it is
missing.  For `last_review`, we'll replace it with a new variable 
`days_since_review`.  Each non-missing observation is formatted YYYY-MM-DD 
so this will be easy to do.  Furthermore we can replace the missing values with
-1 so that the new variable is prepped for modeling.

```{r}
# handle missing `reviews_per_month`
dataset <- dataset %>%
  mutate(reviews_per_month = replace_na(reviews_per_month, 0))
```

```{r}
# handle missing `last_review` by creating `days_since_review` with -1 standing
# in for missing values
dataset <- dataset %>%
  mutate(
    last_review = as.Date(last_review, "%Y-%m-%d"),
    latest_review_day = max(last_review, na.rm = TRUE)
  ) %>%
  mutate(days_since_review = as.numeric(difftime(latest_review_day, last_review, units = "days")) + 1) %>%
  mutate(days_since_review = replace_na(days_since_review, -1)) %>%
  select(-c(last_review, latest_review_day))
```

The other missing observations are in the `host_name` and `name` variables.  
Since there is already a `host_id` field, these are not needed for identifying
distinct hosts.  If we were to use these variables for modeling purposes, we
could replace their missing values with `host_id`.  However, using `name` opens
up privacy and discrimination concerns that we should avoid.
```{r}
# drop host_name and name as protected information
dataset <- select(dataset, -c(name, host_name))
```

## Exploring and visualizating data.

The bar plots below give further insight into our discrete (categorical)
variables.  Our `neighbourhood_group` observations are limited to 5 neighborhood 
groups, with Manhattan and Brooklyn comprising the majority.  Room type is mostly 
evenly split between "entire home/apt" and "private room", with "shared room" 
rarely appearing.  Unsurprisingly, `days_since_review` has too many distinct 
categories to plot - it should probably be treated as a continuous variable.
Finally, there are 221 distinct neighborhoods within our neighborhood groups.
```{r}
DataExplorer::plot_bar(dataset)
```

```{r}
# neighborhood-specific room_type occurences
dataset %>%
  count(neighbourhood_group, room_type) %>%
  arrange(room_type, desc(n))
```

```{r}
# count of distinct neighbourhoods per group
dataset %>%
  distinct(neighbourhood_group, neighbourhood) %>%
  count(neighbourhood_group)
```


Let's further investigate our continuous variables via histograms:
```{r}
plot_histogram(dataset)
```


** are smaller host_id observations correlated with higher average reviews?


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
